---
title: "Test Design coverage of DS sim v2"
author: "Michelle DePrenger-Levin"
date: "2026-02-24"
output: html_document
---

```{r}
library(dssd)
library(dsims)
library(mgcv)

```


The shapefile is not included to protect the location of the recently delisted rare plant
```{r}
picnic.region <- make.region(region.name = "Picnic", shape = ".",
                            units = "m")
plot(picnic.region)

coverP <- make.coverage(picnic.region, n.grid.points = 1000)
```

Make survey == one random realization of a design    
covered area are the rectangles covered by the line transects (2* half-width or truncation)
```{r}

# density surface in relative terms. maintains relative density when given absolute values later 
# Test how much bias will be introduced by non-uniform coverage
# x and y's are removed to protect the location of the population
pdens <- make.density(picnic.region, 
                      x.space = 1, 
                      constant = 1) # relative density

pdens <- add.hotspot(object = pdens,
                     centre = c(x1, y1),
                     sigma = 60,
                     amplitude = 4)

pdens <- add.hotspot(object = pdens,
                     centre = c(x2, y2),
                     sigma = 10,
                     amplitude = 2)

pdens <- add.hotspot(object = pdens,
                     centre = c(x3, y3),
                     sigma = 10,
                     amplitude = 2)

pdens <- add.hotspot(object = pdens,
                     centre = c(x4, y4),
                     sigma = 30,
                     amplitude = -1)


plot(pdens)
pdensity.data <- pdens@density.surface[[1]]

## s() is a smooth term
fit.pgam <- gam(density ~ s(x,y), data = pdensity.data, family = gaussian(link="log"))
pgam.density <- make.density(region = picnic.region,
                            x.space = 1,
                            fitted.model = fit.pgam)

plot(pgam.density)
```

```{r}
## 1247 individuals at Picnic

trnc <- 3
scaleparam <- 1.16
shapeparam <- 2.58 
N <- trunc(1247) # listed as size of cluster but assume talks about size of population
reps = 999
# spacing <- 20# 25# 15# 20
desnAn <- 55

# covariates <- list(size = list(distribution = "ztruncpois", mean = 2)) ## Cactus can occur in groups but often as 1  
# Define the population description
pop.desc <- make.population.description(region = picnic.region,
                                        density = pgam.density,
                                        covariates = list(),
                                        N = N,
                                        fixed.N = TRUE)

# Covariate parameters on the log scale
detect <- make.detectability(key.function = "hr",
                             scale.param = scaleparam,   # 1.16
                             shape.param = shapeparam,   # 2.58
                             # cov.param = cov.param,
                             truncation = trnc)
plot(detect, pop.desc)


## hazard rate with no covariates, hazard rate with cluster size as a covariate
# with default error variance estimator R2: random line placement, unequal line lengths
#                                       R3: random, model-assisted estimator, true contagion process
#                                       S2: systematic, variances weighted by line length
analyses <- make.ds.analysis(dfmodel = list(~1), # list(~1, ~size),
                             key = c("hr"), 
                             truncation = trnc,
                             er.var = "R2",  
                             criteria = "AIC")
###########################################################################################################
```

Coverage scores   
```{r}


for(linelength in c(30*5,30*7.5,30*10,30*12.5,30*15)){ 

###########################################################################################################
## Parallel to gradient
picnic.design <- make.design(region = picnic.region,
                             design = "systematic",
                             line.length = linelength,  ## sets spacing to get close to the total length across samplers
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn,
                             truncation = trnc,
                             coverage.grid = coverP)


## Generate one random set of transects
transects <- generate.transects(picnic.design)

## Assess coverage uniformity and design stats   
# Want uniform coverage across the majority of the survey region
design <- run.coverage(picnic.design, reps = 1000)

#######################################################################################################################################
##################################################
# Perpendicular to gradient, parallel to gradient contours
picnic.design.perp <- make.design(region = picnic.region,
                             design = "systematic",
                             line.length = linelength,  ## sets spacing to get close to the total length across samplers
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn+90,
                             truncation = trnc,
                             coverage.grid = coverP)


## Generate one random set of transects
transects.perp <- generate.transects(picnic.design.perp)

## Assess coverage uniformity and design stats   
# Want uniform coverage across the majority of the survey region
design.perp <- run.coverage(picnic.design.perp, reps = 1000)

##########################################################################
## Segments
## Parallel to gradient
picnic.design.seg <- make.design(region = picnic.region,
                             design = "segmentedgrid",
                             line.length = linelength,  ## sets spacing to get close to the total length across samplers
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn,
                             truncation = trnc,
                             seg.length = 30,
                             coverage.grid = coverP)


## Generate one random set of transects
transects.seg <- generate.transects(picnic.design.seg)

design.seg <- run.coverage(picnic.design.seg, reps = 1000)

####################################################
# Perpendicular to gradient
picnic.design.seg.perp <- make.design(region = picnic.region,
                             design = "segmentedgrid",
                             line.length = linelength,  ## sets spacing to get close to the total length across samplers
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn+90,
                             truncation = trnc,
                             spacing = spacing,
                             seg.length = 30,
                             coverage.grid = coverP)


## Generate one random set of transects
transects.seg.perp <- generate.transects(picnic.design.seg.perp)

design.seg.perp <- run.coverage(picnic.design.seg.perp, reps = 1000)
}



```



Simulation where the effort is kept constant  
Run simulations for multiple population sizes    
Effort as how many 30 meter segments are needed     
```{r}

summarySpacingOrintationType <- do.call(rbind,lapply(c(30*5,30*7.5,30*10,30*12.5,30*15), function(lineLength){ 
  do.call(rbind, lapply(trunc(c(trunc(N/2),N, (N*2))), function(Nseq){
    
    ## 1247 individuals at Picnic

trnc <- 3
scaleparam <- 1.16
shapeparam <- 2.58 
reps = 1000
desnAn <- 55

# covariates <- list(size = list(distribution = "ztruncpois", mean = 2)) ## Cactus can occur in groups but often as 1  
# Define the population description
pop.desc <- make.population.description(region = picnic.region,
                                        density = pgam.density,
                                        covariates = list(),
                                        N = Nseq,
                                        fixed.N = TRUE)

# Covariate parameters on the log scale - what does this refer to???
# cov.param <- list(size = log(1.08))
detect <- make.detectability(key.function = "hr",
                             scale.param = scaleparam,   # 1.16
                             shape.param = shapeparam,   # 2.58
                             # cov.param = cov.param,
                             truncation = trnc)

## hazard rate with no covariates, hazard rate with cluster size as a covariate
# with default error variance estimator R2: random line placement, unequal line lengths
#                                       R3: random, model-assisted estimator, true contagion process
#                                       S2: systematic, variances weighted by line length
analyses <- make.ds.analysis(dfmodel = list(~1),# list(~1, ~size),
                             key = c("hr"),  # ,"hr"),
                             truncation = trnc,
                             er.var = "R2",  # "S2": not supported, likely will be over-estimated, O2 also not supported
                             criteria = "AIC")

##############################################################################################################################
## Parallel to gradient
picnic.design <- make.design(region = picnic.region,
                             design = "systematic",
                             line.length = lineLength,  ## or spacing, line length sets spacing to get close to the total length of all samplers
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn,
                             truncation = trnc,
                             coverage.grid = coverP)


## Generate one random set of transects
transects <- generate.transects(picnic.design)

sim.parallel <- make.simulation(reps = reps,
                                design = picnic.design,
                                population.description = pop.desc,
                                detectability = detect,
                                ds.analysis = analyses)

sim.parallel <- run.simulation(sim.parallel, run.parallel = TRUE)
summarySim <- summary(sim.parallel, use.max.reps = TRUE)
est.df <- as.data.frame(do.call(rbind,lapply(1:dim(sim.parallel@results$individuals$N)[3],function(x){
  out <- sim.parallel@results$individuals$N[,,x]
  out
  }))  )


est.df <- est.df %>%
  summarise(meanEst = mean(Estimate),
         meanLCL = mean(lcl),
         meanUCL = mean(ucl),
         meanCV = mean(cv)) 

parallel.df <- data.frame(parORseg = "systematic",              # systematic grid or segments  
                     perpORparallel = "parallel",   # To the gradient
                     # spacing = spacing,
                     Effort = lineLength, 
                     designAngle = desnAn,
                     popsz = Nseq,
                     numSamplers = transects@samp.count,
                     coverageAvg = transects@cov.area/transects@strata.area,
                     summarySim@individuals$summary,
                     summarySim@individuals$N,
                     summarySim@individuals$D,
                     est.df)

#######################################################################################################################################
# Perpendicular to gradient, parallel to gradient contours
picnic.design.perp <- make.design(region = picnic.region,
                             design = "systematic",
                             line.length = lineLength,  ## or spacing, line length sets spacing to get close to the total length of all samplers
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn+90,
                             truncation = trnc,
                             coverage.grid = coverP)


## Generate one random set of transects
transects.perp <- generate.transects(picnic.design.perp)

sim.parallel.perp <- make.simulation(reps = reps,
                                design = picnic.design.perp,
                                population.description = pop.desc,
                                detectability = detect,
                                ds.analysis = analyses)
eg.parallel.perp <- run.survey(sim.parallel.perp)

sim.parallel.perp <- run.simulation(sim.parallel.perp, run.parallel = TRUE)
summarySimPerp <- summary(sim.parallel.perp, use.max.reps = TRUE)
estperp.df <- as.data.frame(do.call(rbind,lapply(1:dim(sim.parallel.perp@results$individuals$N)[3],function(x){
  out <- sim.parallel.perp@results$individuals$N[,,x]
  out
}))  )

estperp.df <- estperp.df %>%
  summarise(meanEst = mean(Estimate),
         meanLCL = mean(lcl),
         meanUCL = mean(ucl),
         meanCV = mean(cv)) 

perp.df <- data.frame(parORseg = "systematic",              # parallel grid or segments  
                     perpORparallel = "perpendicular",   # To the gradient
                     Effort = lineLength, 
                     designAngle = desnAn+90,
                     popsz = Nseq,
                     numSamplers = transects.perp@samp.count,
                     coverageAvg = transects.perp@cov.area/transects.perp@strata.area,
                     summarySimPerp@individuals$summary,
                     summarySimPerp@individuals$N,
                     summarySimPerp@individuals$D,
                     estperp.df )

##########################################################################
## Segments
## Parallel to gradient
picnic.design.seg <- make.design(region = picnic.region,
                             design = "segmentedgrid",
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn,
                             truncation = trnc,
                             line.length = lineLength, 
                             seg.length = 30,
                             coverage.grid = coverP)


## Generate one random set of transects
transects.seg <- generate.transects(picnic.design.seg)

sim.parallel.seg <- make.simulation(reps = reps,
                                design = picnic.design.seg,
                                population.description = pop.desc,
                                detectability = detect,
                                ds.analysis = analyses)
eg.parallel.seg <- run.survey(sim.parallel.seg)

sim.parallel.seg <- run.simulation(sim.parallel.seg, run.parallel = TRUE)
summarySeg <- summary(sim.parallel.seg, use.max.reps = TRUE)
estseg.df <- as.data.frame(do.call(rbind,lapply(1:dim(sim.parallel.seg@results$individuals$N)[3],function(x){
  out <- sim.parallel.seg@results$individuals$N[,,x]
  out
}))  )

estseg.df <- estseg.df %>%
  summarise(meanEst = mean(Estimate),
         meanLCL = mean(lcl),
         meanUCL = mean(ucl),
         meanCV = mean(cv)) 

seg.df <- data.frame(parORseg = "segments",              # parallel grid or segments  
                     perpORparallel = "parallel",   # To the gradient
                     Effort = lineLength, 
                     designAngle = desnAn,
                     popsz = Nseq,
                     numSamplers = transects.seg@samp.count,
                     coverageAvg = transects.seg@cov.area/transects.seg@strata.area,
                     summarySeg@individuals$summary,
                     summarySeg@individuals$N,
                     summarySeg@individuals$D,
                     estseg.df)


####################################################
# Perpendicular to gradient
picnic.design.seg.perp <- make.design(region = picnic.region,
                             design = "segmentedgrid",
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn+90,
                             truncation = trnc,
                             line.length = lineLength, 
                             seg.length = 30,
                             coverage.grid = coverP)


## Generate one random set of transects
transects.seg.perp <- generate.transects(picnic.design.seg.perp)

sim.parallel.seg.perp <- make.simulation(reps = reps,
                                design = picnic.design.seg.perp,
                                population.description = pop.desc,
                                detectability = detect,
                                ds.analysis = analyses)
eg.parallel.seg.perp <- run.survey(sim.parallel.seg.perp)

sim.parallel.seg.perp <- run.simulation(sim.parallel.seg.perp, run.parallel = TRUE)
summarySegPerp <- summary(sim.parallel.seg.perp, use.max.reps = TRUE)
estsegperp.df <- as.data.frame(do.call(rbind,lapply(1:dim(sim.parallel.seg.perp@results$individuals$N)[3],function(x){
  out <- sim.parallel.seg.perp@results$individuals$N[,,x]
  out
}))  )

estsegperp.df <- estsegperp.df %>%
  filter(!is.na(Estimate)) %>%
  summarise(meanEst = mean(Estimate),
         meanLCL = mean(lcl),
         meanUCL = mean(ucl),
         meanCV = mean(cv)) 

segperp.df <- data.frame(parORseg = "segments",              # parallel grid or segments  
                     perpORparallel = "perpendicular",   # To the gradient
                     Effort = lineLength, 
                     designAngle = desnAn+90,
                     popsz = Nseq,
                     numSamplers = transects.seg.perp@samp.count,
                     coverageAvg = transects.seg.perp@cov.area/transects.seg.perp@strata.area,
                     summarySegPerp@individuals$summary,
                     summarySegPerp@individuals$N,
                     summarySegPerp@individuals$D,
                     estsegperp.df)

dfout <-  parallel.df %>%
  bind_rows(perp.df) %>%
  bind_rows(seg.df) %>%
  bind_rows(segperp.df)

dfout


}))
  
})) ## Systematic and specify the desired total sampler length instead of spacing between samplers

```


## Number of detections for each design   
Simulation where the effort is kept constant  

Run simulations for multiple population sizes    

```{r}
N <- 1247

samplingBYdesign <- do.call(rbind,lapply(c(30*5,30*7.5,30*10,30*12.5,30*15), function(lineLength){ 
  do.call(rbind, lapply(trunc(c(trunc(N/2),N, (N*2))), function(Nseq){
    
        ## 1247 individuals at Picnic
    
    trnc <- 3
    scaleparam <- 1.16
    shapeparam <- 2.58 
    reps = 1000
    desnAn <- 55
    
    # covariates <- list(size = list(distribution = "ztruncpois", mean = 2)) ## Cactus can occur in groups but often as 1  
    # Define the population description
    pop.desc <- make.population.description(region = picnic.region,
                                            density = pgam.density,
                                            covariates = list(),
                                            N = Nseq,
                                            fixed.N = TRUE)
    
    # Covariate parameters on the log scale - what does this refer to???
    # cov.param <- list(size = log(1.08))
    detect <- make.detectability(key.function = "hr",
                                 scale.param = scaleparam,   # 1.16
                                 shape.param = shapeparam,   # 2.58
                                 # cov.param = cov.param,
                                 truncation = trnc)
    
    ## hazard rate with no covariates, hazard rate with cluster size as a covariate
    # with default error variance estimator R2: random line placement, unequal line lengths
    #                                       R3: random, model-assisted estimator, true contagion process
    #                                       S2: systematic, variances weighted by line length
    analyses <- make.ds.analysis(dfmodel = list(~1),# list(~1, ~size),
                                 key = c("hr"),  # ,"hr"),
                                 truncation = trnc,
                                 er.var = "R2",  # "S2": not supported, likely will be over-estimated, O2 also not supported
                                 criteria = "AIC")
    
    ##############################################################################################################################
    ## Parallel to gradient
    picnic.design <- make.design(region = picnic.region,
                                 design = "systematic",
                                 line.length = lineLength,  ## or spacing, line length sets spacing to get close to the total length of all samplers
                                 edge.protocol = "minus",      # don't go outside study area
                                 design.angle = desnAn,
                                 truncation = trnc,
                                 coverage.grid = coverP)
    
    ### repeat 
    systematic.parallel <- do.call(rbind, lapply(1:100, function(i) {
      
      ## Generate one random set of transects
      transects <- generate.transects(picnic.design)
      
      sim.parallel <- make.simulation(reps = reps,
                                      design = picnic.design,
                                      population.description = pop.desc,
                                      detectability = detect,
                                      ds.analysis = analyses)
      eg.parallel <- run.survey(sim.parallel)
      
      parallel.df <- data.frame(parORseg = "systematic",              # systematic grid or segments  
                           perpORparallel = "parallel",   # To the gradient
                           Effort = lineLength, 
                           designAngle = desnAn,
                           popsz = Nseq,
                           numSamplers = transects@samp.count,
                           LineLength = transects@line.length,
                           Detections = length(eg.parallel@dists.in.covered),
                           coverageAvg = transects@cov.area/transects@strata.area)
      
      parallel.df
    
    }))
    
    
    #######################################################################################################################################
    ##################################################
    # Perpendicular to gradient, paralell to gradient contours
    picnic.design.perp <- make.design(region = picnic.region,
                                 design = "systematic",
                                 line.length = lineLength,  ## or spacing, line length sets spacing to get close to the total length of all samplers
                                 edge.protocol = "minus",      # don't go outside study area
                                 design.angle = desnAn+90,
                                 truncation = trnc,
                                 coverage.grid = coverP)
    
    
    ### repeat 
    systematic.perp1 <- do.call(rbind, lapply(1:100, function(i) {
      
      ## Generate one random set of transects
      transects <- generate.transects(picnic.design.perp)
      
      sim.parallel.perp1 <- make.simulation(reps = reps,
                                      design = picnic.design.perp,
                                      population.description = pop.desc,
                                      detectability = detect,
                                      ds.analysis = analyses)
      eg.parallel.perp <- run.survey(sim.parallel.perp1)
      
      parallel.df <- data.frame(parORseg = "systematic",              # systematic grid or segments  
                           perpORparallel = "perpendicular",   # To the gradient
                           Effort = lineLength, 
                           designAngle = desnAn+90,
                           popsz = Nseq,
                           numSamplers = transects@samp.count,
                           LineLength = transects@line.length,
                           Detections = length(eg.parallel.perp@dists.in.covered),
                           coverageAvg = transects@cov.area/transects@strata.area)
      
      parallel.df
    
    }))
    ##########################################################################
    ## Segments
    ## Parallel to gradient
    picnic.design.seg <- make.design(region = picnic.region,
                                 design = "segmentedgrid",
                                 # line.length = ,  ## or spacing, line length sets spacing to get close to the total length of all samplers
                                 edge.protocol = "minus",      # don't go outside study area
                                 design.angle = desnAn,
                                 truncation = trnc,
                                 line.length = lineLength, 
                                 seg.length = 30,
                                 coverage.grid = coverP)
    
    
    ### repeat 
    segment <- do.call(rbind, lapply(1:100, function(i) {
      
      ## Generate one random set of transects
      transects <- generate.transects(picnic.design.seg)
      
      sim.segment <- make.simulation(reps = reps,
                                     design = picnic.design.seg,
                                     population.description = pop.desc,
                                     detectability = detect,
                                     ds.analysis = analyses)
      eg.segment <- run.survey(sim.segment)
      
      parallel.df <- data.frame(parORseg = "segments",              # systematic grid or segments  
                           perpORparallel = "parallel",   # To the gradient
                           Effort = lineLength, 
                           designAngle = desnAn,
                           popsz = Nseq,
                           numSamplers = transects@samp.count,
                           LineLength = transects@line.length,
                           Detections = length(eg.segment@dists.in.covered),
                           coverageAvg = transects@cov.area/transects@strata.area)
      
      parallel.df
    
    }))
    
    ####################################################
    # Perpendicular to gradient
    picnic.design.seg.perp <- make.design(region = picnic.region,
                                 design = "segmentedgrid",
                                 # line.length = ,  ## or spacing, line length sets spacing to get close to the total length of all samplers
                                 edge.protocol = "minus",      # don't go outside study area
                                 design.angle = desnAn+90,
                                 truncation = trnc,
                                 line.length = lineLength, 
                                 seg.length = 30,
                                 coverage.grid = coverP)
    
    
    ### repeat 
    segment.perp <- do.call(rbind, lapply(1:100, function(i) {
      
      ## Generate one random set of transects
      transects <- generate.transects(picnic.design.seg.perp)
      
      sim.segment.perp <- make.simulation(reps = reps,
                                          design = picnic.design.seg.perp,
                                          population.description = pop.desc,
                                          detectability = detect,
                                          ds.analysis = analyses)
      eg.segment <- run.survey(sim.segment.perp)
      
      parallel.df <- data.frame(parORseg = "segments",              # systematic grid or segments  
                           perpORparallel = "perpendicular",   # To the gradient
                           Effort = lineLength, 
                           designAngle = desnAn+90,
                           popsz = Nseq,
                           numSamplers = transects@samp.count,
                           LineLength = transects@line.length,
                           Detections = length(eg.segment@dists.in.covered),
                           coverageAvg = transects@cov.area/transects@strata.area)
      
      parallel.df
    
    }))

  outdf <- do.call(rbind, list(systematic.parallel, systematic.perp1, segment,segment.perp))
  outdf
  }))
})) ## Systematic and specify the desired line length instead of the spacing

```




