---
title: "Density gradients and patches"
author: "Michelle DePrenger-Levin"
date: "2026-09-30"
output: html_document
---

Code for Appendix B   
Evaluating Distance Sampling as a Tool for Monitoring Populations of Rare Plants (Cactaceae: Sclerocactus glaucus and Sclerocactus dawsoniae)    
Michelle DePrenger-Levin, Mary Strecker, Jonathan Amegin      

```{r}

library(dssd)
library(dsims)
library(mgcv)
library(sp)
library(ggplot2)
library(tidyverse)
library(patchwork)
```


Generalize the impact of a density gradient and patchy distribution  
Make a box the size of Picnic   
```{r}

# Target: 7806.717
BoxM <- sqrt(7806.717)
x_coords <- c(0,0,BoxM,BoxM,0)
y_coords <- c(0,BoxM,BoxM,0,0)

poly1 <- sp::Polygon(cbind(x_coords, y_coords))
poly2 <- sp::Polygons(list(poly1), ID = "A")
polySpatial <- sp::SpatialPolygons(list(poly2), proj4string = CRS("epsg:32630"))  

shapefile(x = polySpatial, file = "Data/generic.region.shp")

generic.region <- make.region(region.name = "box", shape = "Data/generic.region.shp")

coverG <- make.coverage(generic.region, n.grid.points = 1000)
```

Make survey == one random realization of a design    
covered area are the rectangles covered by the line transects (2* half-width or truncation)
```{r}

# density surface in relative terms. maintains relative density when given absolute values later 

set.seed(1234)
## Assumption that density is consistent
# Density surface 1
gdens <- make.density(generic.region, 
                      x.space = 1, 
                      constant = 1) 

plot(gdens)
###########################################################
####  Gradients

## 2: plants often do not occupy all suitable habitat
gdens2 <- make.density(generic.region, 
                      x.space = 1, 
                      constant = 0.001) # relative density

gdens2 <- add.hotspot(object = gdens2,
                     centre = c(BoxM - BoxM/4, (BoxM*3)/4),
                     sigma = BoxM/4,
                     amplitude = 2)

gdens2 <- add.hotspot(object = gdens2,
                     centre = c(BoxM - BoxM/4, (BoxM*2)/4),
                     sigma = BoxM/4,
                     amplitude = 2)

gdens2 <- add.hotspot(object = gdens2,
                     centre = c(BoxM - BoxM/4, BoxM/4),
                     sigma = BoxM/4,
                     amplitude = 2)

plot(gdens2) 

##################################################################
## 3: occupy more of the suitable habitat but lower density at habitat edges

gdens3 <- make.density(generic.region, 
                      x.space = 1, 
                      constant = 0.001) # relative density

gdens3 <- add.hotspot(object = gdens3,
                     centre = c(BoxM - BoxM/5, (BoxM*4)/5),
                     sigma = BoxM/2,
                     amplitude = 2)

gdens3 <- add.hotspot(object = gdens3,
                     centre = c(BoxM - BoxM/5, (BoxM*2.5)/5),
                     sigma = BoxM/5,
                     amplitude = 2)

gdens3 <- add.hotspot(object = gdens3,
                     centre = c(BoxM - BoxM/5, BoxM/5),
                     sigma = BoxM/2,
                     amplitude = 2)

plot(gdens3) 
##################################################################
## 4: random intensity of gradient
set.seed(4)
sigmaRan <- runif(9, BoxM/5,BoxM/2)
set.seed(4)
ampRan <- rpois(9,2)

gdens4 <- make.density(generic.region, 
                      x.space = 1, 
                      constant = 0.001) # relative density

gdens4 <- add.hotspot(object = gdens4,
                     centre = c(BoxM - BoxM/5, (BoxM*4)/5),
                     sigma = sigmaRan[1],
                     amplitude = ampRan[1])

gdens4 <- add.hotspot(object = gdens4,
                     centre = c(BoxM - BoxM/5, (BoxM*2.5)/5),
                     sigma = sigmaRan[2],
                     amplitude = ampRan[2])

gdens4 <- add.hotspot(object = gdens4,
                     centre = c(BoxM - BoxM/5, BoxM/5),
                     sigma = sigmaRan[3],
                     amplitude = ampRan[3])

plot(gdens4) 

## 5
gdens5 <- make.density(generic.region, 
                      x.space = 1, 
                      constant = 0.001) # relative density

gdens5 <- add.hotspot(object = gdens5,
                     centre = c(BoxM - BoxM/5, (BoxM*4)/5),
                     sigma = sigmaRan[4],
                     amplitude = ampRan[4])

gdens5 <- add.hotspot(object = gdens5,
                     centre = c(BoxM - BoxM/5, (BoxM*2.5)/5),
                     sigma = sigmaRan[5],
                     amplitude = ampRan[5])

gdens5 <- add.hotspot(object = gdens5,
                     centre = c(BoxM - BoxM/5, BoxM/5),
                     sigma = sigmaRan[6],
                     amplitude = ampRan[6])

plot(gdens5)

## 6
gdens6 <- make.density(generic.region, 
                      x.space = 1, 
                      constant = 0.001) # relative density

gdens6 <- add.hotspot(object = gdens6,
                     centre = c(BoxM - BoxM/5, (BoxM*4)/5),
                     sigma = sigmaRan[7],
                     amplitude = ampRan[7])

gdens6 <- add.hotspot(object = gdens6,
                     centre = c(BoxM - BoxM/5, (BoxM*2.5)/5),
                     sigma = sigmaRan[8],
                     amplitude = ampRan[8])

gdens6 <- add.hotspot(object = gdens6,
                     centre = c(BoxM - BoxM/5, BoxM/5),
                     sigma = sigmaRan[9],
                     amplitude = ampRan[9])

plot(gdens6)
##################################################################
## 7: consistent gradient of density across all suitable habitat

# random slopes
set.seed(5)
slopes <- runif(5, 0.1, 6)

gdens7 <- make.density(generic.region, 
                       x.space = 1,
                       y.space = 1,
                       density.formula = paste("0.0001 +", slopes[2], "*x") )

plot(gdens7)

gdens8 <- make.density(generic.region, 
                       x.space = 1,
                       y.space = 1,
                       density.formula = paste("0.0001 +", slopes[3], "*x") )

plot(gdens8)

gdens9 <- make.density(generic.region, 
                       x.space = 1,
                       y.space = 1,
                       density.formula = paste("0.0001 +", slopes[4], "*x") )

plot(gdens9)

gdens10 <- make.density(generic.region, 
                       x.space = 1,
                       y.space = 1,
                       density.formula = paste("0.0001 +", slopes[5], "*x") )

plot(gdens10)

gdens11 <- make.density(generic.region, 
                       x.space = 1,
                       y.space = 1,
                       density.formula = paste("0.0001 +", slopes[1], "*x") )

plot(gdens11)

gdens1_11 <- list(gdens,gdens2,gdens3,gdens4,gdens5,gdens6,gdens7,gdens8,gdens9,gdens10,gdens11)

##################################################################
#### Patches (12-21)

## 12-21: uneven suitability within a population and short dispersal distances cause patchy distribution. There are gradients at smaller scales but not consistent across the entire population

gdens12 <- make.density(generic.region, 
                      x.space = 1, 
                      constant = 0.0001) 

gdens12_21 <- lapply(1:10, function(i){
  for(hots in 1:rpois(1,36)){
    gdens12 <- add.hotspot(object = gdens12,
                           centre = c(runif(1,0,BoxM),runif(1,0,BoxM)),
                           sigma = rpois(1,5),
                           amplitude = rpois(1,2))
  }
  gdens12
  })


plot(gdens12_21[[2]])

#########################################################################
## There can be a general density gradient and patches within

## 22-31: start with density surface 2 and add random hotspots

gdens22_31 <- lapply(1:10, function(i){
  for(hots in 1:rpois(1,36)){
    gdens2 <- add.hotspot(object = gdens2,
                          centre = c(runif(1,0,BoxM),runif(1,0,BoxM)),
                          sigma = rpois(1,5),
                          amplitude = rpois(1,2))
  }
  gdens2
  })


plot(gdens22_31[[5]])

densitySurfaces <- do.call(c, list(gdens1_11, gdens12_21, gdens22_31))

save(densitySurfaces, file = "Data/densitySurfaces.rdata")

```

